A tool set for random number generation on GPUs in R

### 2022-07-22 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.Rnw
typo p1
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3123 edb9625f-4e0d-4859-8d74-9fd3b1da38cb
update
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3181 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

Now we use the 4 streams created in section \ref{createstreams} to generate a vector of double-precision i.i.d. $U(0,1)$ random numbers with \fct{runifGpu}. To view the generated random numbers, we need to convert them to \proglang{R} vectors or matrices, by doing this, the random numbers are moved from the global memory of the GPU device to the host.
<<simStuff2, eval=TRUE,tidy=TRUE, echo=TRUE>>=
sim_1 =  runifGpu(n = 8, streams = myStreamsGpu, Nglobal = c(2,2))
as.vector(sim_1)
@
The arguments of the \fct{runifGpu} are described as follows:
\begin{itemize}
\itemsep0em 
  \item \code{n}: Either a scalar giving the number of samples to generate or a vector of length 2 specifying the size of a matrix to be output.
  \item \code{streams}: Streams for random number generation, which must be stored on the GPU as a \code{vclMatrix} object.
  \item \code{Nglobal}: Number of work items, by default it is set as $(64,8)$.
  \item \code{type}: \code{double}, \code{float} or \code{integer}, the format of generated random numbers, defaults to \code{double}.
  \item \code{verbose}: Print extra information if $\text{verbose} > 1$.
\end{itemize}

@@

 <<Timecomparemonth,echo=FALSE,eval=TRUE, cache=FALSE, results="hide">>=
 # datamonth<-read.csv("/home/ruoyong/diseasemapping/pkg/gpuRandom/inst/documents/paper1_2021_5/month.csv")
 # month<-as.matrix(datamonth[,-1])
 # rownames(month) <- c("Jan", "Feb","Mar","Apr","May","Jun","Jul","Aug","Sept","Oct","Nov","Dec")
 # colnames(month) <- c("Ane", "Men", "Cya", "Her", "Omp", "Gas", "Lim", "Cle", "Pal", "Dow", "Chr", "Hyp")
 month <- as.matrix(readRDS(system.file("extdata", "month.Rds", package = "clrng")))
 @
 <<echo=FALSE,eval=TRUE>>= 
sn <- function(x,digits)
{
  if (x==0) return("0")
  ord <- floor(log(abs(x),10))
  x <- x / 10^ord
  if (!missing(digits)) x <- format(x,digits=digits)
  if (ord==0) return(as.character(x))
  return(paste(x,"\\times 10^{",ord,"}",sep=""))
}
@
 <<monthdata,eval=TRUE,echo=FALSE, cache=TRUE, fig.pos='h', message=FALSE,purl=FALSE>>=
library(knitr)
library(tidyverse)
library(kableExtra)
knitr::kable(month,format="latex", align = c("rrrrrrrrrrrr"), #label="tab:month"
  caption = "Monthly counts of birth anomalies.\\label{tab:month}", booktabs=TRUE,linesep = "") %>%
  kable_styling(full_width = F, position = "center")#latex_options = "HOLD_position")
@

@@

Table \ref{tab:week} is a $7 \times 12$ table shows frequencies for congenital anomalies of the newborn by birth day of week in 2018 within the United States. 
<<Timecompareweek, echo=FALSE, eval=TRUE, cache=TRUE,results="hide", fig.pos='h'>>=
# dataweek<-read.csv("/home/ruoyong/diseasemapping/pkg/gpuRandom/inst/documents/paper1_2021_5/weekday.csv")
# week<-as.matrix(dataweek[,-1])
# rownames(week) <- c("Mon", "Tue","Wed","Thu","Fri","Sat","Sun")
# colnames(week) <- c("Ane", "Men", "Cya", "Her", "Omp", "Gas", "Lim", "Cle", "Pal", "Dow", "Chr", "Hyp")
week <- as.matrix(readRDS(system.file("extdata", "week.Rds", package = "clrng")))
@
<<weekdata, eval=TRUE, echo=FALSE, cache=TRUE>>=
knitr::kable(week,format="latex",align = c("rrrrrrrrrrrr"),  #label = "tab:week",
             caption = "Day-of-week birth anomaly data\\label{tab:week}", booktabs=TRUE,linesep = "") %>%
kable_styling(full_width = F, position = "center")#, latex_options = "HOLD_position")
@
<<TimecompareweekGpu2, eval=TRUE, cache=FALSE>>=
# using GPU
week_GPU<-gpuR::vclMatrix(week,type="integer")
system.time(resultWeek<-clrng::fisher.sim(week_GPU, 1e7, streams=streams,
                type="double",returnStatistics=TRUE,Nglobal = c(256,64)))
unlist(resultWeek[c('threshold','simNum','counts')])
resultWeek$p.value
@
<<TimecompareweekGpu3, echo=FALSE, eval=TRUE, cache=TRUE>>=
resultWeek$cpu = as.vector(resultWeek$sim)
@
<<TimecompareweekCpu, cache=TRUE, eval=TRUE>>=
# using CPU
system.time(result_weekcpu<-fisher.test(week,simulate.p.value = TRUE,B=10010624))
result_weekcpu$p.value
@
The ``week'' table has a much smaller p-value: around $\Sexpr{sn(result_weekcpu$p.value,3)}$, which should require a larger number of simulations to get a more accurate p-value. With more than ten million simulations, we get 1350 cases and a p-value around $\Sexpr{sn(resultWeek$p.value,3)}$ with \fct{fisher.sim}. \fct{stats::fisher.test} takes about 91.8 seconds, while \fct{fisher.sim} takes about 1.9 seconds, the GPU elapsed time is decreased to about 2.1\% of the time taken on CPU.

>>>>

Now we use the 4 streams created in section \ref{createstreams} to generate a vector of double-precision i.i.d. $U(0,1)$ random numbers with \fct{runifGpu}. To view the generated random numbers, we need to convert them to \proglang{R} vectors or matrices, by doing this, the random numbers are moved from the global memory of the GPU device to the host.
<<simStuff2, eval=TRUE,tidy=TRUE, echo=TRUE>>=
sim_1 =  runifGpu(n = 8, streams = myStreamsGpu, Nglobal = c(2,2))
as.vector(sim_1)
@
The arguments of the \fct{runifGpu} are described as follows:
\begin{itemize}
\itemsep0em 
  \item \code{n}: Either a scalar giving the number of samples to generate or a vector of length 2 specifying the size of a matrix to be output.
  \item \code{streams}: Streams for random number generation, which must be stored on the GPU as a \code{vclMatrix} object.
  \item \code{Nglobal}: Number of work items, by default it is set as $(64,8)$.
  \item \code{type}: \code{double}, \code{float} or \code{integer}, the format of generated random numbers, defaults to \code{double}.
  \item \code{verbose}: Print extra information if $\text{verbose} > 1$, default is set to 0.
\end{itemize}

@@

 <<Timecomparemonth,echo=FALSE,eval=TRUE, cache=FALSE, results="hide">>=
 # datamonth<-read.csv("/home/ruoyong/diseasemapping/pkg/gpuRandom/inst/documents/paper1_2021_5/month.csv")
 # month<-as.matrix(datamonth[,-1])
 # rownames(month) <- c("Jan", "Feb","Mar","Apr","May","Jun","Jul","Aug","Sept","Oct","Nov","Dec")
 # colnames(month) <- c("Ane", "Men", "Cya", "Her", "Omp", "Gas", "Lim", "Cle", "Pal", "Dow", "Chr", "Hyp")
 #month <- as.matrix(readRDS(system.file("extdata", "month.Rds", package = "clrng")))
 data("month")
 month <- as.matrix(month)
 @
 <<echo=FALSE,eval=TRUE>>= 
sn <- function(x,digits)
{
  if (x==0) return("0")
  ord <- floor(log(abs(x),10))
  x <- x / 10^ord
  if (!missing(digits)) x <- format(x,digits=digits)
  if (ord==0) return(as.character(x))
  return(paste(x,"\\times 10^{",ord,"}",sep=""))
}
@
 <<monthdata,eval=TRUE,echo=FALSE, cache=TRUE, fig.pos='h', message=FALSE,purl=FALSE>>=
library(knitr)
library(tidyverse)
library(kableExtra)
knitr::kable(month,format="latex", align = c("rrrrrrrrrrrr"), #label="tab:month"
  caption = "Monthly counts of birth anomalies.\\label{tab:month}", booktabs=TRUE,linesep = "") %>%
  kable_styling(full_width = F, position = "center")#latex_options = "HOLD_position")
@

@@

Table \ref{tab:week} is a $7 \times 12$ table shows frequencies for congenital anomalies of the newborn by birth day of week in 2018 within the United States. 
<<Timecompareweek, echo=FALSE, eval=TRUE, cache=TRUE,results="hide", fig.pos='h'>>=
# dataweek<-read.csv("/home/ruoyong/diseasemapping/pkg/gpuRandom/inst/documents/paper1_2021_5/weekday.csv")
# week<-as.matrix(dataweek[,-1])
# rownames(week) <- c("Mon", "Tue","Wed","Thu","Fri","Sat","Sun")
# colnames(week) <- c("Ane", "Men", "Cya", "Her", "Omp", "Gas", "Lim", "Cle", "Pal", "Dow", "Chr", "Hyp")
#week <- as.matrix(readRDS(system.file("extdata", "week.Rds", package = "clrng")))
data("week")
as.matrix(week)
@
<<weekdata, eval=TRUE, echo=FALSE, cache=TRUE>>=
knitr::kable(week,format="latex",align = c("rrrrrrrrrrrr"),  #label = "tab:week",
             caption = "Day-of-week birth anomaly data\\label{tab:week}", booktabs=TRUE,linesep = "") %>%
kable_styling(full_width = F, position = "center")#, latex_options = "HOLD_position")
@
<<TimecompareweekGpu2, eval=TRUE, cache=FALSE>>=
# using GPU
week_GPU<-gpuR::vclMatrix(week,type="integer")
system.time(resultWeek<-clrng::fisher.sim(week_GPU, 1e7, streams=streams,
                type="double",returnStatistics=TRUE,Nglobal = c(256,64)))
unlist(resultWeek[c('threshold','simNum','counts')])
resultWeek$p.value
@
<<TimecompareweekGpu3, echo=FALSE, eval=TRUE, cache=TRUE>>=
resultWeek$cpu = as.vector(resultWeek$sim)
@
<<TimecompareweekCpu, cache=TRUE, eval=TRUE>>=
# using CPU
system.time(result_weekcpu<-fisher.test(week,simulate.p.value = TRUE,B=10010624))
result_weekcpu$p.value
@
The ``week'' table has a much smaller p-value: around $\Sexpr{sn(result_weekcpu$p.value,3)}$, which should require a larger number of simulations to get a more accurate p-value. With more than ten million simulations, we get 1350 cases and a p-value around $\Sexpr{sn(resultWeek$p.value,3)}$ with \fct{fisher.sim}. \fct{stats::fisher.test} takes about 91.8 seconds, while \fct{fisher.sim} takes about 1.9 seconds, the GPU elapsed time is decreased to about 2.1\% of the time taken on CPU.
